---
title: "Word Frequency Analysis Notebook"
output: 
  html_notebook:
    theme: cerulean
    toc: yes
    toc_float: yes
---

# set options {.tabset}

## disable scientific notation
```{r}
options(scipen=999)
```
## show only 2 decimals
```{r}
options(digits = 2)
```

# Load packages {.tabset}

## Tidyverse
Tidyverse contains many functions useful to cleaning, tidying, and manipulating data

Documentation can be viewed [here](https://www.tidyverse.org)

```{r}
if (!require(tidyverse)){
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse)
}
```
## tidytext
tidytext makes it easy to mine text data

Documentation can be viewed [here](https://www.tidytextmining.com)

```{r}
if (!require(tidytext)){
  install.packages("tidytext", dependencies = TRUE)
  library(tidytext)
}
```

# Import data {.tabset}

## get list of private school documents

```{r}
(docs.private <- list.files("Private/"))
```
## get list of public school documents

```{r}
(docs.public <- list.files("Public/"))
```

## import private school alerts
```{r}
for (doc in docs.private)
{
  if(!exists("alerts.private"))
  {
    enframe(read_lines(paste("Private/", doc, sep=""))) %>%
      mutate(name = doc,
             institutionType = "Private") -> alerts.private
  }
  else
  {
    enframe(read_lines(paste("Private/", doc, sep=""))) %>%
      mutate (name = doc,
              institutionType = "Private") %>%
      bind_rows(alerts.private) -> alerts.private
  }
}
alerts.private
```
## import public school alerts
```{r}
for (doc in docs.public)
{
  if(!exists("alerts.public"))
  {
    enframe(read_lines(paste("Public/", doc, sep=""))) %>%
      mutate(name = doc,
             institutionType = "Public")-> alerts.public
  }
  else
  {
    enframe(read_lines(paste("Public/", doc, sep=""))) %>%
      mutate(name = doc,
             institutionType = "Public") %>%
      bind_rows(alerts.public) -> alerts.public
  }
}
alerts.public
```

## combine private and public alerts
```{r}
(alerts.private %>%
  bind_rows(alerts.public)-> alerts.all)
```

# tidy data {.tabset}

```{r}
(alerts.all %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, value) %>%
  filter(!str_detect(word, "[:digit:]")) %>%
  anti_join(stop_words) %>%
  filter(word != "ut")-> alerts.all.tidy)
```

# Word frequencies {.tabset}


## Most frequent words

```{r echo=FALSE}

num_rows <- nrow(alerts.all.tidy)

alerts.all.tidy %>%
  count(word, sort = TRUE) %>%
  mutate(n_rows = num_rows,
         proportion = n/n_rows) %>%
  top_n(10, wt = n) %>%
  ggplot(aes(x = reorder(word, proportion), label = sprintf("%.1f %%", 100*proportion), y = proportion)) + 
  geom_col(alpha = 0.8, show.legend = FALSE) +
  geom_text(hjust = 0, nudge_y = 0.0005, size = 3) +
  coord_flip() +
  labs(x = "",
       y = "Proportion of times word used across all alerts") + 
  ggtitle("Most frequent words across all crime alerts") +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA))
```

## Frequency of keywords
```{r echo=FALSE}
num_rows <- nrow(alerts.all.tidy)

alerts.all.tidy %>%
  count(word, sort = TRUE) %>%
  mutate(n_rows = num_rows,
         proportion = n/n_rows) %>%
  filter(word == "subject" | word == "student" | word == "suspect" | word == "victim" | word == "male" | word == "criminal" | word == "black" | word == "white" | word == "sexual" | word == "intoxication" | word == "unknown" | word == "crime" | word == "alcohol" | word == "use" | word == "person") %>%
  ggplot(aes(x = reorder(word, proportion), label = sprintf("%.1f %%", 100*proportion), y = proportion)) + 
  geom_col(alpha = 0.8, show.legend = FALSE) +
  geom_text(hjust = 0, nudge_y = 0.0005, size = 3) +
  coord_flip() +
  labs(x = "",
       y = "Proportion of times key words used across all alerts") + 
  ggtitle("Frequency of keywords across all crime alerts") +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA))
```






















